Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-08-01 emsoft-2019/camera_ready/Cont_Syn.tex
Took out the paragraph on compkexity.
updates

\begin{figure}
\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,mathescape=true,basicstyle=\small\ttfamily,escapeinside={<@}{@>}]
 model = input() 
 $\Delta$ = input()
 $\varepsilon_{Q}$ = input()
 assert ($\Delta$ >= 0 and $\varepsilon_{Q}$ > 0)
 $\varepsilon$ = $\varepsilon_{A/D}$ + $\varepsilon_{Q}$<@$\quad$//\,\,width of the tube@>

>>>>

\begin{figure}
\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,mathescape=true,basicstyle=\small\ttfamily,escapeinside={<@}{@>}]
 model = input() 
 $\Delta$ = input()
 $\varepsilon_{Q}$ = input()
 assert ($\Delta$ >= 0 and $\varepsilon_{Q}$ > 0)
 <@\textcolor{blue}{$\varepsilon$ = $\max_i \{||F_i||_2\}\varepsilon_{A/D}+\varepsilon_{Q}$}@>$\quad$//width of the tube

### 2019-08-01 emsoft-2019/camera_ready/Error_Analysis.tex
updates

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{equation}
  \max\{\|v_{i}-v_{j}\|,\,\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} +  \max\{\mathit{err}(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{equation}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions 
$\mathcal R_i,\mathcal R_j$ (two regions are neighboring if their intersection is non-empty).
The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x_k$ \RM{in $\mathcal{R}_i \cap \mathcal{R}_j$? Or all states?}. 
The term $\mathit{err}(v_{i})_{p}$ in
the second part captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

% Note that $\mathcal R_i\subset \bar{\mathcal R}_i$. In other words, $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$.
% Note that $\bar{\mathcal R}_i\cap \bar{\mathcal R}_j$ gives us the states where one of the two regions $\mathcal R_i, \mathcal R_i$ might be selected instead of the other one because of the fixed-precision implementation of the regions.   
% $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$ that include all states in $\mathcal R_i$this definition expands the regions $\mathcal $
 %That is, with infinite-precision arithmetic
 Therefore, the first term in~\autoref{eq:delta} can be computed less conservatively by maximizing only over 
 the expanded borders:
 \begin{equation}
 \label{eq:maximization}
 \max_{x,i,j}\{\|F_ix+G_i-F_jx-G_j\|,\,x\in \bar{\mathcal B}_{ij}\}.
% \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|
 \end{equation}

@@

\subsection{Approximate Control Output}

>>>>

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{equation}
  \max\{\|v_{i}-v_{j}\|,\,\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} +  \max\{\mathit{err}(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{equation}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions 
$\mathcal R_i,\mathcal R_j$ (two regions are neighboring if their intersection is non-empty).
The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x$ %\RM{
in $\mathcal{R}_i \cap \mathcal{R}_j$. %}
The term $\mathit{err}(v_{i})_{p}$ in
the second part captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

% Note that $\mathcal R_i\subset \bar{\mathcal R}_i$. In other words, $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$.
% Note that $\bar{\mathcal R}_i\cap \bar{\mathcal R}_j$ gives us the states where one of the two regions $\mathcal R_i, \mathcal R_i$ might be selected instead of the other one because of the fixed-precision implementation of the regions.   
% $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$ that include all states in $\mathcal R_i$this definition expands the regions $\mathcal $
 %That is, with infinite-precision arithmetic
 Therefore, the first term in~\autoref{eq:delta} can be computed less conservatively by maximizing only over 
 the expanded borders:
 \begin{equation}
 \label{eq:maximization}
 \max_{x,i,j}\{\|F_ix+G_i-F_jx-G_j\|,\,x\in \bar{\mathcal B}_{ij}\}.
% \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|
 \end{equation}
 %%%%%%%%%%%%%%
 	The first part of the error in~\autoref{eq:delta} captures the error due to
 	selecting of the incorrect region.
 	This would happen since the matrices $H_i,K_i$ are stored in the hardware with fixed-point formats. In order to quantify the error, we define the expanded border $\bar{\mathcal B}_{ij}$ as the tube around the border between the regions $\mathcal{R}_i$ and $\mathcal{R}_j$:
 	\begin{equation*}
 	\bar{\mathcal B}_{ij} := \left\{x\in\mathbb R^n\,|\, \| H_i x - K_i\| \le \varepsilon\,\, \text{ and }\,\, \| H_j x - K_j\| \le \varepsilon \right\},
 	\end{equation*}
 	where $\varepsilon$ captures the uncertainty in computing the correct region.
 	%
 	The width of the tube $\varepsilon$ is due to two sources of errors:
 	\Mahmoud{\begin{itemize}
 		\item analog-to-digital conversion $\varepsilon_{A/D}$: using the quantized output of the analog-to-digital converter $\hat x$ instead of the actual state $x$. We assume $\|x - \hat x\|\le \varepsilon_{A/D}$.
 		\item quantization of region
 		bounds in memory $\varepsilon_{Q}$: using $\hat H_i,\hat K_i$ which are the quantized versions of $H_i,K_i$ with $p$ bits precision. This is related to the boundaries of the regions.
 		%\item using $\hat F_i,\hat G_i$ which are the quantized versions of $F_i,G_i$ with $p$ bits precision. This is related to the affine functions.
 	Note that the error resulted from using the quantized versions of $F_i,G_i$ is discussed in \autoref{subsec:app_cont_err}.
 	 %We denote to controllers $\kappa(\mathcal H,x)$ and $\kappa(\mathcal{\hat H},\hat)$ to distinguish between the case that fixed point representation is used or not.
 	We define the fixed point realization of the controller in \autoref{eq:affine_map} as:
 	 	\begin{equation}
 	 	\label{eq:affine_map_hat}
 	 	\hat\kappa(\hat x_k)=
 	 	\begin{cases}
 	 	\hat F_1\hat x_k+\hat G_1 & \text{if $\hat x_k\in \mathcal{\hat R}_1$}\\
 	 	\hat F_2\hat x_k+\hat G_2 & \text{if $\hat x_k\in \mathcal{\hat R}_2$}\\
 	 	\vdots\\
 	 	\hat F_P x_k+\hat G_P & \text{if $ \hat x_k\in \mathcal{\hat R}_P$}
 	 	\end{cases} 
 	 	\end{equation}
 	 	where $\mathcal{\hat R}_i$ is a polyhedral region determined by a set of linear inequalities $\mathcal{\hat R}_i= \{ \hat x\in\mathcal X\,|\, \hat H_i\hat x\leq \hat K_i\}$. Let us also define the sets $\mathcal H := \{H_i,K_i,i=1,2,\ldots, P\}$ and $\mathcal F:=\{F_i,G_i,i=1,2,\ldots, P\}$. Similarly, we define $\hat{\mathcal H}:= \{\hat H_i,\hat K_i,i=1,2,\ldots, P\}$ and $\hat{\mathcal F}:=\{\hat F_i,\hat G_i,i=1,2,\ldots, P\}$. Using the triangle inequality we have:
 	 	\|\kappa(x)-\hat \kappa(\hat x)
 	 	where $\kappa(\mathcal H^*,\mathcal F^*,x^*)$ denotes the controller defined over the sets $\mathcal F^*$, $\mathcal H^*$ and the state $x^*$. The second term corresponds to the control approximation error and will be discussed in \autoref{subsec:app_cont_err}.
 	 	Therefore, we can compute the tube width $\varepsilon$ as:
 	 	\begin{equation}\label{eq:epsilontot}
 	 	\varepsilon=\max_i \{||F_i||_2\}\varepsilon_{A/D}+\varepsilon_{Q}
 	 	\end{equation}
% 	\Mahmoud{We define the fixed point realization of the controller in \autoref{eq:affine_map} as:
% 	\begin{equation}
% 	\label{eq:affine_map_hat}
% 	\hat\kappa(\hat x_k)=
% 	\begin{cases}
% 	\hat F_1\hat x_k+\hat G_1 & \text{if $\hat x_k\in \mathcal{\hat R}_1$}\\
% 	\hat F_2\hat x_k+\hat G_2 & \text{if $\hat x_k\in \mathcal{\hat R}_2$}\\
% 	\vdots\\
% 	\hat F_P x_k+\hat G_P & \text{if $ \hat x_k\in \mathcal{\hat R}_P$}
% 	\end{cases} 
% 	\end{equation}
% 	where $\hat x_k$ denotes the output of the analog-to-digital converter corresponding to the input $x_k$. Further, $\hat F_i, \hat G_i$ denote the fixed point representations corresponding to $F_i, G_i$ and $\mathcal{\hat R}_i$ is a polyhedral region determined by a set of linear inequalities $\mathcal{\hat R}_i= \{ \hat x\in\mathcal X\,|\, \hat H_i\hat x\leq \hat K_i\}$.}
%
%	\Mahmoud{Let us define the sets $\mathcal H := \{H_i,K_i,i=1,2,\ldots, P\}$ and $\mathcal F:=\{F_i,G_i,i=1,2,\ldots, P\}$. Similarly, we define $\hat{\mathcal H}:= \{\hat H_i,\hat K_i,i=1,2,\ldots, P\}$ and $\hat{\mathcal F}:=\{\hat F_i,\hat G_i,i=1,2,\ldots, P\}$, where $\hat F_i$, $\hat G_i$, $\hat H_i$ and $\hat K_i$ denote the fixed point representations corresponding to $F_i$, $G_i$, $H_i$ and $K_i$, respectively. }
%
% 	\begin{equation}\label{eq:epsilontot1}
% 	\varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
% 	\end{equation}
 	%
 	%
 	Analog conversion happens just before the controller receives the sensor input from the plant
 	and is given by:
 	\begin{equation*}
 	\varepsilon_{A/D}=\frac{V_{cc}}{2^{r}-1}
 	\end{equation*}
 	where $V_{cc}$ is the reference voltage of the converter (e.g. typically 5V), $r$
 	is the number of bits available to quantize the analog signal, and $2^{r}$ is
 	the resolution of the converter.\\
 	While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
 	$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
 	\begin{equation}\label{eq:quantizationlines}
 	\varepsilon_{Q} \ge  \|(H_i - \hat{H}_i) \hat{x} + (K_i - \hat{K}_i)\|, \quad \forall \hat{x},\forall i,
 	\end{equation}
 	where $\hat{x}$ is the finite-precision value of $x$.
 	%\eva{For which index is this evaluated, $i$, $j$ or both?}
 	That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
 	precision ($H$ and $K$ matrices), and its counter-part quantized by $p$ bits 
 	($\hat{H}_{p}$ and $\hat{K}_{p}$).
 	This second error can be tuned providing a trade-off between accuracy and
 	memory storage required.
 	Therefore, the first term in~\autoref{eq:delta} can be computed less conservatively by maximizing only over 
 	the expanded borders:
 	\begin{equation}
 	\label{eq:maximization}
 	\max_{x,i,j}\{\|F_ix+G_i-F_jx-G_j\|,\,x\in \bar{\mathcal B}_{ij}\}.
 	% \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|
 	\end{equation}

@@

\subsection{Approximate Control Output}
\label{subsec:app_cont_err}

### 2019-08-01 emsoft-2019/camera_ready/Motiv.tex
updates

		%\centering
		%\hspace*{\fill} 
		\begin{tikzpicture}[auto,node distance=1cm,>=latex']
		\node [block,text width=3.8cm, pin={[pinstyle]above:$\Delta=\Delta_0$},
		node distance=1cm, align=center, label={[anchor=south east, draw=black, inner sep=1pt]south east:MATLAB}] (RMPC) {Robust explicit MPC\\with disturbance set $\mathcal W_{\Delta}$};
		\node [block1,below = 0.8cm of RMPC] (highprec) {High-precision controller (\autoref{eq:affine_map})};
		%\node[] (partitions) at (2.5,.5){\includegraphics[width=.08\textwidth]{Figs/regs.jpg}};
		\node [block3,below=.8cm of highprec,
		node distance=1cm,text width=4.4cm,  label={[anchor=south east, draw=black, inner sep=1pt]south east:MATLAB}] (Uij) {Incorrect region selection error};
		\node [draw, diamond, minimum height=3em, minimum width=3em, below =0.8cm of Uij, node distance=1cm] (decide) {$e\leq  \Delta$};
		\node [block3, below=0.8cm of decide,node distance=1cm, label={[anchor=south east, draw=black, inner sep=1pt]south east:Daisy}] (done) {Mixed-precision implementation};
		\node [block2, right=1.3cm of decide,
		node distance=1cm] (delta) {$\Delta=e+\varepsilon_{SAFE}$};
		\node [block1,below = 0.8cm of done] (output) {Low-precision controller (\autoref{eq:affine_map})};
		%\node [scale=1] at (3,-.25)  {$u(x)=Kx+v$};
		%\node at (9,-.25) {$e$};

@@

We can implement the controller using uniform $32$ bit integers, or by selecting a uniform bitwidth.
These take, respectively, $XX$ and $YY$ bits.
\RM{how many bits do we need for the uniform prec impl? how many bits for 32 bits?} 
But we can do better.
We use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, for this example, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses 32 bits.

>>>>

		%\centering
		%\hspace*{\fill} 
		\begin{tikzpicture}[auto,node distance=1cm,>=latex']
		\node [block,text width=3.8cm, pin={[pinstyle]above:$\Delta=\Delta_0$},
		node distance=1cm, align=center, label={[anchor=south east, draw=black, inner sep=1pt]south east:MATLAB}] (RMPC) {Robust explicit MPC\\with disturbance set $\mathcal W_{\Delta}$};
		\node [block1,below = 0.8cm of RMPC] (highprec) {High-precision controller (\autoref{eq:affine_map})};
		%\node[] (partitions) at (2.5,.5){\includegraphics[width=.08\textwidth]{Figs/regs.jpg}};
		\node [block3,below=.8cm of highprec,
		node distance=1cm,text width=4.4cm,  label={[anchor=south east, draw=black, inner sep=1pt]south east:MATLAB}] (Uij) {Incorrect region selection error};
		\node [draw, diamond, minimum height=3em, minimum width=3em, below =0.8cm of Uij, node distance=1cm] (decide) {$e\leq  \Delta$};
		\node [block3, below=0.8cm of decide,node distance=1cm, label={[anchor=south east, draw=black, inner sep=1pt]south east:Daisy}] (done) {Mixed-precision implementation};
		\node [block2, right=1.3cm of decide,
		node distance=1cm] (delta) {$\Delta=e+\varepsilon_{SAFE}$};
		\node [block1,below = 0.8cm of done] (output) {Low-precision controller (\Mahmoud{\autoref{eq:affine_map_hat}})};
		%\node [scale=1] at (3,-.25)  {$u(x)=Kx+v$};
		%\node at (9,-.25) {$e$};

@@

We can implement the controller using uniform $32$ bit integers, or by selecting a uniform bitwidth.
These take, respectively, $13824$ and $6744$ bits.
%\RM{how many bits do we need for the uniform prec impl? how many bits for 32 bits?} 
But we can do better.
We use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, for this example, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses 32 bits.
